DRAFT: Go Implementation of Techniques for Weighted Language Equivalence

### 2020-10-07 paper/paper.tex
updated tests with useful parameters

\title{DRAFT: Go Implementation of Up-To Techniques for Weighted Language Equivalence}

@@

\begin{abstract}
Weighted automata generalize non-deterministic automata by adding
a quantity expressing the weight (or probability) of the execution of each transition.
In this work we propose an implementation of two algorithms for computing language 
equivalence in finite state weighted automata (WAs). The first, a
linear partition refinement algorithm, computes the largest linear weighted bisimulation
for any given LWA (Linear Weighted Automaton) through an iterative method, 
the second algorithm checks the language equivalence 
of two vectors (states) for a given weighted automata by using an additional
data structure representing a congruence relation.
We then compare results of the two algorithms to verify their correctness
and performance on randomly generated samples. 
We finally provide a comparison and runtime statistics.
\end{abstract}

>>>>

\title{DRAFT: Go Implementation of Techniques for Weighted Language Equivalence}

@@

\begin{abstract}
Weighted automata generalize non-deterministic automata by adding
a quantity expressing the weight (or probability) of the execution of each transition.
In this work we propose an implementation of two algorithms for computing
the language equivalence relation in finite state weighted automata (WAs). 
The first algorithm checks the language equivalence 
of two vectors (states) in a weighted automaton, 
with an up-to technique by using an additional
data structure representing a congruence relation.
The second algorithm, a
linear partition refinement algorithm, computes the 
whole equivalence relation, precisely the
largest linear weighted bisimulation,
by linearizing the state space and iteratively refining
the relation. 
We then compare results of the two algorithms to verify their correctness
and performance on randomly generated samples. 
We finally provide a comparison and runtime statistics.
\end{abstract}

### 2020-10-07 paper/sections/algorithms.tex
updated tests with useful parameters

\section{Algorithms}
The first algorithm we implement to compute language equivalence, called \texttt{HKC},
is adapted from \cite{DBLP:journals/corr/Bonchi0K17}. 
The algorithm returns \texttt{true} $\iff \sqmap{v_1} - \sqmap{v_2}$.
It was first introduced by 
Bonchi and Pous in \cite{bonchi2013checking}.
The algorithm, extending the Hopcroft and Karp procedure 
\cite{hopcroft1971linear} with \textit{congruence closure}, is 
proven to be sound and complete \cite{DBLP:journals/corr/Bonchi0K17}.
It is defined for 
WAs over semirings, but in this implementation we are only 
considering fields, in particular 
the field of real numbers ($\K = \R$).
The problem of checking language equivalence 
has been proven undecidable for an arbitrary semiring, so termination 
may not always be guaranteed. However, it has been shown to be decidable
for a broad range of semirings, for example, all the complete and
distributive lattices.
\texttt{HKC} computes $v_1 \sim_l v_2$ for a given weighted automaton
$W = (X, t, o)$ and two vectors $v_1, v_2 \in \K^X$. 
by computing a congruence closure,
and it does so without linearizing the state space.

@@

% TODO
(TODO chiedi a filippo: nelle conclusioni di \cite{BONCHI201277} dice che è diverso 
dall'algoritmo visto nel paper di boreale \cite{boreale2009weighted}. è il caso 
anche per i field e i numeri reali???? o sono lo equivalenti?).

Adapted from \cite{BONCHI201277}, 
\texttt{BPR} is a fixed-point iterative method that, given an LWA
$L = (V, t, o)$,
it computes a basis of the subspace of $V$
representing the complementary relation of $\llwb$ (we later show it to be the 
orthogonal complement in case $V$ is an inner product space). 
Another version of the algorithm is defined in the same work,
called \textit{Forward Partition Refinement}, which directly computes
a basis for $\llwb$
but is shown to be way less efficient than the backwards version.

>>>>

\section{Algorithms}
The first algorithm we implement to compute language equivalence, called \texttt{HKC},
is adapted from \cite{DBLP:journals/corr/Bonchi0K17}. 
The algorithm returns \texttt{true} $\iff \sqmap{v_1} = \sqmap{v_2}$.
It was first introduced by 
Bonchi and Pous in \cite{bonchi2013checking}.
The algorithm, extending the Hopcroft and Karp procedure 
\cite{hopcroft1971linear} with \textit{congruence closure}, is 
proven to be sound and complete \cite{DBLP:journals/corr/Bonchi0K17}.
It is defined for 
WAs over semirings, but in this implementation we are only 
considering fields, in particular 
the field of real numbers ($\K = \R$).
The problem of checking language equivalence 
has been proven undecidable for an arbitrary semiring, so termination 
may not always be guaranteed. However, it has been shown to be decidable
for a broad range of semirings, for example, all the complete and
distributive lattices.
\texttt{HKC} computes $v_1 \sim_l v_2$ for a given weighted automaton
$W = (X, t, o)$ and two vectors $v_1, v_2 \in \K^X$. 
by computing a congruence closure,
and it does so without linearizing the state space.

@@

Adapted from \cite{BONCHI201277}, 
\texttt{BPR}, or \textit{Backwards Partition Refinement}
is a fixed-point iterative method that, given an LWA
$L = (V, t, o)$, it computes the whole

computes a basis of the subspace of $V$
representing the complementary relation of $\llwb$ (we later show it to be the 
orthogonal complement in case $V$ is an inner product space). 
Another version of the algorithm is defined in the same work,
called \textit{Forward Partition Refinement}, which directly computes
a basis for $\llwb$ but is shown to be way less efficient than the backwards version.
The advantage of the backwards
It has been first introduced in \cite{boreale2009weighted}.
Our implementation is directly modeled on the algorithm shown in 
\cite{boreale2009weighted}, since we are fixing the field 
to $\R$ and computing the orthogonal complements instead of 
dual spaces and annihilators. In \cite{BONCHI201277}, the authors 
extended.
